---
title: "Exploratory Data Analysis"
author: "Valarie Bryant"
date: "2/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries, Data
```{r }
#importing libraries related to time series, visualization
library(TSstudio)
library(ggplot2)
library(fpp)
library(tidyverse)
library(astsa)
library(xts)
```

```{r}
#reading in the merged dataset
# data merging done in Python Visualization file
data <- read_csv('MergedTS_Feb17.csv')
data$Date <- as.Date(paste(data$Date,"-01",sep=""))
data <- data %>%
  select(-X1)
head(data)
```
# Basic Time Series
```{r}
#reading in the data as time series, monthly
ts_unemploy = ts(data$unemployment, frequency = 12)

ts_sentiment = ts(data$consumer_sentiment, frequency = 12)
ts_sentiment = na.remove(ts_sentiment)

ts_g_unemploy = ts(data$g_unemployment, frequency = 12)
ts_g_unemploy = na.remove(ts_g_unemploy)

ts_g_food_stamps = ts(data$g_food_stamps, frequency = 12)
ts_g_food_stamps = na.remove(ts_g_food_stamps)

ts_g_gas = ts(data$g_gas, frequency = 12)
ts_g_gas = na.remove(ts_g_gas)

ts_food = ts(data$fs_households, frequency = 12)
ts_food = na.remove(ts_food)

ts_gas = ts(data$retail_gas, frequency = 12)
ts_gas <- na.remove(ts_gas)


```



# Lagged Plots, Consumer Sentiment
```{r}
p <- gglagplot(ts_sentiment, do.lines=FALSE) +
  ggtitle('Lag Plot of Consumer Sentiment')
p
ggsave('lag_sentiment.png')
```

#Decomposed Unemployment Rate
```{r}
jpeg("decompose_unemploy.jpg", width = 1100, height = 800)

decompose_unemploy = decompose(ts_unemploy, "additive")

plot(decompose_unemploy)
dev.off()
```

# ACF Plot, PACF, and Dickey-Fuller Test of food stamps

```{r}
# https://stackoverflow.com/questions/43151069/acf-plot-title-change/51893811
jpeg("acf_food.jpg", width = 1100, height = 400, quality = 100)

acf(ts_food, lag.max = 48, main = '48 Month Lag on Food Stamps')
dev.off()
```

```{r}
# https://stackoverflow.com/questions/43151069/acf-plot-title-change/51893811
jpeg("pacf_food.jpg", width = 1100, height = 400, quality = 100)

pacf(ts_food, lag.max = 48, main = '48 Month PACF on Food Stamps')
dev.off()
```

```{r}
#df.ts=ts(WWWusage)
autoplot(ts_food)
tseries::adf.test(ts_food)

#P-value is greater than 0.05. So obviously the series in not stationary.
# If the P-value is greater than .05, the series isn't stationary

```


# ACF Plot, PACF, and Dickey-Fuller Test of gas prices

```{r}
# https://stackoverflow.com/questions/43151069/acf-plot-title-change/51893811
jpeg("acf_gas.jpg", width = 1100, height = 400, quality = 100)

acf(ts_gas, lag.max = 24, main = '24 Month Lag on Gas Prices')
dev.off()
```

```{r}
# https://stackoverflow.com/questions/43151069/acf-plot-title-change/51893811
jpeg("pacf_gas.jpg", width = 1100, height = 400, quality = 100)

pacf(ts_gas, lag.max = 24, main = '24 Month PACF on Gas Prices')
dev.off()
```


# ADF test on gas prices
```{r}
#df.ts=ts(WWWusage)
autoplot(ts_food)
tseries::adf.test(ts_gas)

#P-value is greater than 0.05. So obviously the series in not stationary.
# If the P-value is greater than .05, the series isn't stationary

```

#google trends for Cheap gas, plotting
```{r}
jpeg("g_gas_difference.jpg", width = 800, height = 800, quality = 300)

par(mfrow = c(3,1))
df1=diff(ts_g_gas)
df2 = diff(df1)
plot(ts_g_gas, main = "Time Series of Google Trends for 'Cheap Gas'")
plot(df1, main ="1st Order Differencing on Google Trends for 'Cheap Gas'")
plot(df2, main ="2nd Order Differencing on Google Trends for 'Cheap Gas'")
dev.off()
```



(c). If the series is not stationary, use a necessary method/s to make the series stationary. Plot the ACF to see it's stationarity.
```{r}
jpeg("g_gas_difference_acf.jpg", width = 800, height = 800, quality = 300)

par(mfrow = c(3,1))
acf(ts_g_gas, main = "ACF of Google Trends for 'Cheap Gas'")
acf(df1, main = "ACF of 1st Differenced Google Trends for 'Cheap Gas" )
acf(df2, main = "ACF of 2nd Differenced Google Trends for 'Cheap Gas")
dev.off()
```

```{r}
jpeg("decompose_gas.jpg", width = 800, height = 600, quality = 300)

ts_gas %>%
  stl(s.window="periodic", robust=TRUE) %>%
  autoplot(main = 'Decomposed US Gas Prices, 1993-2021')
dev.off()
```

# Moving Average Smoothing

#Gas Prices
```{r}
jpeg("MA_Smooth_gas.jpg", width = 1000, height = 600, quality = 300)

ma3g <- autoplot(ts_gas, series="Data") +
  autolayer(ma(ts_gas,3), series="3-MA") +
  xlab("Month") + ylab("Price of Gas ($)") +
  ggtitle("US Retail Gas Prices, 3-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
                      breaks=c("Data","3-MA"))

ma6g <- autoplot(ts_gas, series="Data") +
  autolayer(ma(ts_gas,6), series="6-MA") +
  xlab("Month") + ylab("Price of Gas ($)") +
  ggtitle("US Retail Gas Prices, 6-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","6-MA"="red"),
                      breaks=c("Data","6-MA"))

ma12g <- autoplot(ts_gas, series="Data") +
  autolayer(ma(ts_gas,12), series="12-MA") +
  xlab("Month") + ylab("Price of Gas ($)") +
  ggtitle("US Retail Gas Prices, 12-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","12-MA"="red"),
                      breaks=c("Data","12-MA"))

ma24g <- autoplot(ts_gas, series="Data") +
  autolayer(ma(ts_gas,24), series="24-MA") +
  xlab("Month") + ylab("Price of Gas ($)") +
  ggtitle("US Retail Gas Prices, 24-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","24-MA"="red"),
                      breaks=c("Data","24-MA"))



grid.arrange(ma3g, ma6g,ma12g,ma24g, nrow = 2, ncol=2)
dev.off()

```


#Moving Average Smoothing on Consumer Sentiment
```{r}
jpeg("MA_Smooth_cs.jpg", width = 1300, height = 600, quality = 300)

ma3s <- autoplot(ts_sentiment, series="Data") +
  autolayer(ma(ts_sentiment,3), series="3-MA") +
  xlab("Month") + ylab("Consumer Sentiment") +
  ggtitle("Consumer Sentiment, 3-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","3-MA"="red"),
                      breaks=c("Data","3-MA"))

ma6s <- autoplot(ts_sentiment, series="Data") +
  autolayer(ma(ts_sentiment,6), series="6-MA") +
  xlab("Month") + ylab("Consumer Sentiment") +
  ggtitle("Consumer Sentiment, 6-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","6-MA"="red"),
                      breaks=c("Data","6-MA"))

ma12s <- autoplot(ts_sentiment, series="Data") +
  autolayer(ma(ts_sentiment,12), series="12-MA") +
  xlab("Month") + ylab("Consumer Sentiment") +
  ggtitle("Consumer Sentiment, 12-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","12-MA"="red"),
                      breaks=c("Data","12-MA"))

ma24s <- autoplot(ts_sentiment, series="Data") +
  autolayer(ma(ts_sentiment,24), series="24-MA") +
  xlab("Month") + ylab("Consumer Sentiment") +
  ggtitle("Consumer Sentiment, 24-month smoothing") +
  scale_colour_manual(values=c("Data"="grey50","24-MA"="red"),
                      breaks=c("Data","24-MA"))

grid.arrange(ma3s, ma6s,ma12s,ma24s, nrow = 2, ncol=2)
dev.off()

```