---
title: "Financial Analysis: Target Stock"
author: "Valarie Bryant"
date: "4/20/2021"
output: html_document
---
```{r}
#reading in required libraries
rm(list = ls(all = TRUE))   # Remove objects from the previous session
library(tidyverse)
library(TSA)
library(fGarch) 
library(dynlm)
library(tseries)
library(astsa) 
library(xts)
library(fpp)
library(astsa)
library(ggplot2)
```

data source: 
https://finance.yahoo.com/quote/TGT/history?p=TGT 

```{r}
#reading in data
target <- read_csv("TGT.csv")
head(target)
```


```{r}
#plotting the data
jpeg("Output/TGT_Close.jpg", width = 800, height = 600, quality = 100)

close <- ts(target$Close)
plot(close, main = "Target Closing Stock Price, April 2016- April 2021")
dev.off()
```


```{r}
#differencing, logging, to get returns, and plotting returns.
jpeg("Output/TGT_Returns.jpg", width = 1000, height = 300, quality = 100)

r.target=diff(log(close))*100
plot(r.target, main = "Differenced Log of Target Closing Prices")
dev.off()
```


```{r}
# Examining the PACF of the returns
#jpeg("Output/TGT_Returns_PACF.jpg", width = 800, height = 700, quality = 100)

par(mfrow=c(2,1))

acf(r.target)
pacf(r.target)
#dev.off()
```


```{r}
# ACF & PACF of the squared series
#jpeg("Output/TGT_Returns_sq_PACF.jpg", width = 800, height = 700, quality = 100)
par(mfrow=c(2,1))

acf(r.target^2)
pacf(r.target^2)
#dev.off()
```
The ACF and PACF only have a first value that is barely statistically significant, suggesting we have white noise OR an ARMA(1,1) model.
Looking at the ACF/ PACF of the squared returns, we may want to fit a GARCH(2,2)


```{r}
#fitting an ARIMA model on the returns themselves
arma = Arima(r.target, order=c(1,0,1)) 
arma.re= arma$residuals
```


#running GARCH on raw returns, with no ARMA
```{r}
summary(garchFit(~garch(2,2),r.target, trace = F))

summary(garchFit(~garch(2,1),r.target, trace = F)) #### this is the best, based on AIC/BIC
summary(garchFit(~garch(1,2), r.target, trace = F)) #### this returns lots of NA values, no standard errors?

summary(garchFit(~garch(1,1),r.target, trace = F))

```


# Models with an ARMA(1,1), then GARCH
```{r}
summary(garchFit(~garch(2,2),arma.re,trace = F))
summary(garchFit(~garch(1,1),arma.re,trace = F))
summary(garchFit(~garch(2,1),arma.re,trace = F))
summary(garchFit(~garch(1,2),arma.re,trace = F))
```


```{r}
#looking at those basic residuals
jpeg("Output/TGT_ARMA.jpg", width = 800, height = 500, quality = 100)

checkresiduals(arma.re)
dev.off()
```



```{r}
#examining the residuals of a GARCH (1,1) on the raw returns
jpeg("Output/TGT_GARCH.jpg", width = 800, height = 500, quality = 100)

fit2 = garch(r.target,order=c(1,1),trace=F)
checkresiduals(fit2) #relatively doing a good job
dev.off()

```


```{r}
Box.test(fit2$residuals, type = "Ljung")
```
 We have a high p-value, so we do not reject the null hypothesis of independence, which is good!
